\bf Formatting Your Paper \\ for the \confname~Conference

### 2016-06-10 acl.2016/amta2016.tex
add amta style

\documentclass[]{article}
\usepackage[letterpaper]{geometry}
\usepackage{amta2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{natbib}
\usepackage{layout}

@@

\begin{abstract}
  This document contains the instructions for preparing a camera-ready
  manuscript for the proceedings of \confname . The document itself
  conforms to its own specifications, and is therefore an example of
  what your manuscript should look like. Authors are asked to conform
  to all the directions reported in this document.
\end{abstract}

\section{Credits}

This text format is derived from that of the Journal of Evolutionary
Computation.\footnote{Originally written by Darrell Whitley, and only
  later modified by Marc Schoenauer, especially regarding the
  bibliography style} The instructions in this document are derived
from those of the MT Summit 2013, % update this list at next change
which was the latest in a long line of adaptations: EAMT 2011, 2010
and 2009, Coling 2008, ACL-07, Coling/ACL-06, EACL-06, ACL-05 and
EACL-03.  All these were based on the formats of earlier ACL and EACL
Conference proceedings.  Those versions were written by several
people, including John Chen, Henry S. Thompson and Donald Walker.

@@

The following formatting instructions are directed to authors of
papers accepted for publication in \confname{}~proceedings, including
the main conference, workshops, posters and demos. See also the
conference website \website~for additional advice and information
regarding submission.  All authors are required to adhere to these
specifications. Authors are required to submit their papers in PDF
(Portable Document Format). Papers are designed to be printed on
\textbf{US letter paper}. Authors from countries in which access to
word processing systems is limited should contact the appropriate
program chair(s) as soon as possible.

\section{General Instructions}

Manuscripts must be in single-column format. The {\nobreak{title}}
must be centered at the top of the first page. Authors' names, email
addresses and complete mailing addresses must appear below the title,
as in the current document. {\bf Type single-spaced.}  Start all pages
directly under the top margin. See the guidelines later regarding
formatting the first page.

Unless otherwise specified, the maximum length of a manuscript is
\paperlength~pages, plus 2 (two) pages for references
%%% for regular and workshop papers and \shortpaperlength~pages for posters and demos, 
(see Section~\ref{sec:length} for additional information on the
maximum number of pages).

\subsection{Electronically-available resources}

This description is provided in \LaTeX2e (\nobreak{\conffilename.tex})
along with the \LaTeX2e style file used to format it
(\nobreak{\conffilename.sty}) and bibliography
(\textsc{Bib}\TeX) example file (\nobreak{\conffilename.bib}); and in
PDF format (\nobreak{\conffilename.pdf}). These files are all
available at \downloadsite. There is also a Microsoft Word document
template (\nobreak{\conffilename.dot}) available at the same URL. We
strongly re\-commend the use of these style files, which have been
appropriately tailored for the \confname~proceedings.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). This format can be generated directly
from \LaTeX source files, using program pdf\LaTeX, or it can be
generated from postscript files. On Unix systems, you can use {\tt
  ps2pdf} for this purpose.  Recent versions of Microsoft Windows can
produce PDF directly (File$>$Save As$>$Save As Type: PDF). Otherwise,
you can use Adobe's Distiller or GSview (File$>$Convert$>$pdfwrite);
if you have \textit{cygwin} installed, you can use
\textit{ps2pdf}. Note that some word processing programs generate PDF
which may not include all the necessary fonts (esp. tree diagrams,
symbols). When you print or create the PDF file, there is usually an
option in your printer setup to include none, all or just non-standard
fonts.  Please make sure that you select the option of including ALL
the fonts. {\em Before sending it, test your PDF by printing it from a
  computer different from the one where it was created.} Moreover,
some word processors may generate very large postscript / PDF files,
where each page is rendered as an image. Such images may reproduce
poorly. In this case, try alternative ways to obtain the postscript
and / or PDF. One way on some systems is to install a driver for a
postscript printer, send your document to the printer specifying
``Output to a file'', then convert the file to PDF.

% US Letter
It is important to specify the \textbf{US Letter format} (21.6 cm x
27.9 cm) / (8.5 in x 11 in) when formatting the paper. When working with {\tt
dvips}, for instance, one should specify {\tt -t letter}.

% A4
%It is of utmost importance to specify the \textbf{A4 format} (21.0 cm
%x 29.7 cm) / (8.3 in x 11.7 in) when formatting the paper. When working with
%{\tt dvips}, for instance, one should specify {\tt -t a4}.

Print-outs of the PDF file on US Letter paper should look like the
present document, which conforms to the formatting requirements. {\em
  Note that in order for your paper to print correctly, you should
  disable centering and scale-to-fit options on your printer.} If you
cannot meet the above requirements about the production of your
camera-ready paper, please contact the program chairs as soon as
possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts in a single column, in the manner these
instructions are formatted. The exact dimensions for a page on
US letter paper are:

% US letter
\begin{itemize}
\item Left margin: 4.45 cm (1.75 in)
\item Right margin: 3.8 cm (1.5 in)
\item Top margin: 3.8 cm (1.5 in)
\item Bottom margin: 3.2 cm (1.25 in)
\item Text width: 13.3 cm (5.25 in)
\item Text height: 21.0 cm (8.25 in)
\end{itemize}

Pages should not be numbered. In LaTeX, this can be achieved by
inserting  
\begin{quote}
\begin{verbatim}
\pagestyle{empty}
\end{verbatim}
\end{quote}
after the \verb=\maketitle= command.
% \layout{}

\subsection{Fonts}

For uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use {\bf Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

\subsection{The First Page}
\label{ssec:first}

Draw a horizontal line the full width of the text above the title, and
another one below the authors' names and affiliations.  Center the
title across the top of the page. Authors' names appear under the
title, one per line, with the author name in boldface, flush to the
left and corresponding email flush to the right.  Affiliations and
mail addresses appear under each corresponding author name, flush to
the left. Do not use footnotes for affiliations. Do not include the
paper ID number assigned during the submission process. Use the
single-column format throughout the text.

{\bf Title}: Place the title centered at the top of the first page, in
a 18-point bold font. Long titles should be typed on two lines without
a blank line intervening. Approximately, put the title at 3.8 cm (1.5
in) from the top of the page, followed by a blank line, then the
authors' names in 11-point bold font, and the affiliations in 11-point
regular font on the following line. Do not use only initials for given
names (middle initials are allowed). Avoid capitalizing last
names. The affiliation should contain the author's complete address,
and if possible an electronic mail address. Leave about 1 cm (0.4 in)
between the affiliation and the body of the first page.

{\bf Abstract}: Type the abstract at the beginning of the first
column. The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about 1
cm (0.4 in) on each side. The word {\bf Abstract} should be in 9-point
bold font above the body of the abstract, which should be in 9-point
font. The abstract should be a concise summary of the general thesis
and conclusions of the paper. It should be no longer than 200 words.

{\bf Text}: Begin typing the main body of the text immediately after
the abstract. Use 10 point font for text. {\bf Indent} when starting a
new paragraph, except for the first paragraph following each heading.

\subsection{Sections}

{\bf Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals. Do not number subsubsections. Use 9-point bold
font for subsection headings and 11-point bold font for section
headings.

{\bf Citations}: Citations within the text appear in parentheses
as~\citep{Smith} or, if the author's name appears in the text itself,
as \cite{Smith}. Citations in parentheses should not be used as
linguistic phrases; for example, instead of ``\citep{Smith}
\nobreak{argues} that \ldots'' say ``\cite{Smith} \nobreak{argues}
that \ldots''.  Treat double authors as in~\citep{DD15}, but write as
in~\citep{PDC} when more than two authors are involved.
\nobreak{Append} lowercase letters to the year in cases of ambiguity
as in \citep{JonesFirst}.  Collapse multiple citations in parenthesis
as in~\citep{Smith,JonesFirst} and like this for multiple citations
with the same-named \nobreak{author:}
\citep{Smith,SmithConc,JonesFirst,JonesSecond}
%(Tam and Schultz, 2006, 2007; Gledson and Keane, 2008a,b).

\textbf{References}: Gather the full set of references together under the
heading {\bf References}; place the section before any Appendices, unless they
contain references. Arrange the references alphabetically by the first
author's last-name, rather than by order of occurrence in the text, and invert
the first-name and last-name of the first author (only). Provide as complete
a citation as possible, using a consistent format, such as the one for {\em
North American Opera\/}.  Use of full names for
authors rather than initials is preferred.  Use full names for journals and
conferences, not abbreviations (for example ``45th Meeting of the Association
for Computational Linguistics'', not ``ACL07'').

The \LaTeX2e{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations,
short citations and multiple citations as described above.

{\bf Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\textbf{Acknowledgement} section should go as a last section immediately
\textit{before the references}.  Do not number the acknowledgement section.

\subsection{Footnotes}

{\bf Footnotes}: Put footnotes at the bottom of the page and use
9-point font. They may be numbered or referred to by asterisks or
other symbols.\footnote{This is how a footnote should appear.}
Footnotes should be separated from the main text by a
line.\footnote{Note the line separating the footnotes from the text.}

\subsection{Copyright}

% \confname{} requires a copyright license statement. This should be inserted as
% an unnumbered footnote on the first column of the first page. The \LaTeX{} style file
% (and the Word template) adds copyright statement automatically. Refer to the style file for instructions to
% change this if required.

\confname{} lets copyright stay with the authors and allows for free verbatim distribution of papers for optimal dissemination. A footer with the text
\begin{quote}
\begin{small}
\begin{it}
  Proceedings of the Twelfth Conference of the Association for Machine Translation in the Americas, Austin, Texas, 29 October - 3 November 2016.\\
  Beregovaya, O., Doyon, J., Green, S., Langlois, L., Richardson, S., Schwartz, L., eds.\\
  \copyright{} 2016 The authors.\\
  This article is licensed under a Creative Commons 3.0 license, no derivative works, attribution, CC-BY-ND.
\end{it}
\end{small}
\end{quote}
will be added to the bottom of all papers when proceedings are
edited. If, for some reason, this copyright notice is not adequate for
your paper, please contact the program chairs as soon as possible.

\subsection{Graphics}

{\bf Illustrations}: Place figures, tables, and photo\-graphs in the
paper near where they are first discussed, rather than at the end, if
possible.

{\bf Captions}: Provide a caption for every illustration; number
each one sequentially in the form:  ``Figure 1. Caption of the
Figure.'', ``Table 1. Caption of the Table.''  Type the captions of
the figures and tables below the body, using 10-point text.

\section{Length of Submission}
\label{sec:length}

Unless otherwise specified, the maximum length is \paperlength~pages,
plus 2 (two) pages for references. 
%%%%% for regular and workshop papers. and \shortpaperlength pages for posters and demos. 
The page limit should be observed strictly. All illustrations and
appendices must be accommodated within these page limits, following
the formatting instructions given in the present document.

>>>>

\documentclass[]{article}
\usepackage[letterpaper]{geometry}
\usepackage{amta2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{natbib}
\usepackage{layout}
\usepackage{algorithmic}
\usepackage{graphicx}

@@

\begin{abstract}
The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community. As commercial use has increased, there is need for  software that is optimized for commercial requirements, in particular, fast phrase-based decoding and more efficient utilization of modern multicore servers.

In this paper we re-examine the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability on multicore machines. The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales monotonically with the number of cores. 
\end{abstract}

@@

SMT has steadily progressed from a research discipline to commercial viability during the past decade as can be seen from services such as the Google and Microsoft Translation services. As well as general purpose services such as these, there is a large number of companies that offer customized translation systems, as well as  companies and organization that implement in-house solutions. Many of these customized solutions use Moses as their SMT engine.

For many users, decoding is the most time-critical part of the translation process. Making use of the multiple cores that are now ubiquitous in todays servers is a common strategy to ameliorate this issue. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\cite{mfernandez2016boosting}. That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores. There have been speculation on the causes of the inefficiency as well as potential remedies.

This paper is the first we know of that focuses on improving decoding speed on multicore servers. We will take a holistic approach to solving this issue. We create a decoder that is optimized for multi-core processing speed by concentrating on four main areas.
\begin{enumerate}
  \item \vspace{-2 mm} Faster memory management of data-structures through the use of customized memory pools
  \item \vspace{-2 mm} Exploring alternatives to cardinality-based hypothesis stack configuration
  \item \vspace{-2 mm} Re-examining the efficiency of phrase-table lookup using translation rule caching and data compression
  \item \vspace{-2 mm} Integrating the lexicalized re-ordering model into the phrase-table, thus eliminating the need for independent random lookup this model 
\end{enumerate}
The result is a decoder that is significantly faster than the Moses baseline for single-threaded operation, and scales with the number of cores.

As far as possible, model scores and functionality are compatible with Moses to aid comparison and ease transition for existing users. All source code will be made available under an open-source license.

The rest of the paper will be broken up into the following sections. The rest of this section will discuss prior work and an outline of the phrase-based model. Section~\ref{sec:Proposed Improvements} will then describe the modifications to improve decoding speed. We describe the experiment setup in Section~\ref{sec:Experimental Setup} and present results Section~\ref{sec:Results}. We conclude in the last section and discuss possible future work.

\subsection{Prior Work}

Most prior work on increasing decoding speed look to optimizing specific components of the decoder or the decoding algorithm.

~\citep{Heafield-kenlm} and ~\citep{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models. ~\citep{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements. ~\citep{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.

~\citep{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm which allows the tradeoff between speed and translation quality to the adjusted with a single parameter. ~\citep{wuebker2012fast} note that language model querying is amongst the most expensive operation in decoding. They sought to improved decoding speed by caching score computations early pruning of translation options. This work is similar to ~\citep{Heafield-mtplz} which group hypotheses with identical language model context and incrementally expand them, reducing LM querying.

~\citep{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.

There are a number of phrase-based decoding implementations, many of which implements the extensions to the phrase-based model described above.

The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.

Joshua~\cite{Joshua-Decoder} also supports cube-pruning for phrase-based models. Phrasal~\cite{spence2014phrasal} supports a number of variants of the phrase-based model. Jane~\cite{peitz2012jane} supports the language model look-ahead described in ~\citep{wuebker2012fast} in addition to other tools to speed up decoding such as having a separate fast, lightweight decoder. mtplz is a specialized decoder developed to implement the incremental decoding described in ~\citep{Heafield-mtplz}.

The Moses, Joshua and Phrasal decoders implement multithreading, however, they all report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{\scriptsize https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{\scriptsize https://twitter.com/ApacheJoshua/status/342022794097340416}).

Jane and mtplz are single-threaded decoders, relying on external applications to parallelize operations. %This has the advantage it is easier to parallelize processing across multiple servers, not just multiple cores. As we shall see, this method does not scale beyond a small number of cores when attempting to scale to multiple cores within one server.

This paper not only focuses on faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers.

\subsection{Phrase-Based Model}

The objective of decoding is to find the target translation with the maximum probability, given a source sentence. That is, for a source sentence $s$, the objective is to find a target translation $\hat{t}$ which has the highest conditional probability $p(t | s)$. Formally, this is written as:
\begin{equation}
\hat{t} = \arg \max_t p( t | s )
\label{eq:argmax-factored-trans}
\end{equation}
where the \emph{arg max} function is the search. The log-linear model generalizes Equation~\ref{eq:argmax-factored-trans} to include more component models and weighting each model according to the contribution of each model to the total probability. 
\begin{equation}
\label{eq:Log-Linear}
p(t | s) 	=  \frac{1}{Z} \exp ( \sum_m \lambda_m h_m ( t, s)^{} )
\end{equation}
where $\lambda_m$ is the weight, and $h_m$ is the \emph{feature function}, or `score', for model $m$. $Z$ is the partition function which can be ignored for optimization.

The standard feature functions in the phrase-based model include:
\begin{enumerate}
  \item \vspace{-2 mm} a distortion penalty
  \item \vspace{-2 mm} a phrase-penalty,
  \item \vspace{-2 mm} a word penalty,
  \item \vspace{-2 mm} an unknown word penalty.
  \item \vspace{-2 mm} log transforms of the target language model probability $p(t)$, 
  \item \vspace{-2 mm} log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
  \item \vspace{-2 mm} log transforms of the lexicalized re-ordering probabilities,
\end{enumerate}
%which we have suffixed with $_{TM} $ to avoid confusion with the overall model probability $ p(t|s) $ and $ p(s|t)$.

Of these feature functions, we will focus on optimizing the speed of the phrase-table and lexicalized re-ordering models.

%The first four feature functions are simple and do not leave much room for optimization. As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\citep{Heafield-kenlm} and ~\citep{yasuhara-EtAl:2013:EMNLP}.

\subsection{Beam Search}

A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once. Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis. This \emph{hypothesis expansion} process starts with a hypothesis that has translated no source word and ends with completed hypotheses that has translated all source words. The highest-scoring completed hypothesis, according to the model score, is considered the best translation, $\hat{t} $.

In the phrase-based model, each rule translates a contiguous sequence of source words. Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit. The target output is constructed strictly left-to-right from the target side from the series of translation rules.

A beam search algorithm is used to create the completed hypothesis set efficiently. Hypotheses are grouped into stacks where each stack holds a number of comparable hypotheses. Most phrase-based implementations group hypotheses according to coverage cardinality. % in the same stack that have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,... |s| \} $ of the number of  source words translated.

\section{Proposed Improvements}
\label{sec:Proposed Improvements}

We will also concentrate on four main areas for optimization.

\subsection{Efficient Memory Allocation}

The search algorithm creates and destroy a large number of intermediate objects such as hypotheses and feature function states. The memory management this requires puts a burden on the operating system due to the need to synchronize memory access. This problem worsens with multi-threaded applications such as Moses.

Libraries such as tcmalloc~\cite{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.

We shall seek to improve decoding speed by replacing the operating system's general purpose memory management with our own custom memory management scheme. Memory will be allocated from a memory pool rather than use the operating system's general purpose allocation functions.

A memory pool is a large block of memory that has been given to the application by the operating system. The application is then responsible for allocating portions of this memory to its components as required. We will use thread-specific memory pools to increase speed by avoiding locking contention during memory access. Our memory pools will be dynamic, growing in size when required but never reducing in size. The pools are not deleted until the decoder application ends.

To further increase memory management speed, objects in the memory pool are not deleted. Unused data structures accumulates in the pool until a reset event. After the event, the pool is assumed to be empty and simply reused.

We instantiate two memory pools per decoding thread, one which is never reset and another which is reset after the decoding of each sentence. Data structures are created in either pool according to their life cycle.

Accumulating unused objects in the memory pools can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle unused objects before the reset event. This also increase speed as LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.

%The speed of memory access is dependent on whether the data is in the CPU cache which is a constrained resource compared to memory size, typically 20MB in the latest processors. We shall seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.

%In multiprocessor servers, the CPU cache is attached to each processor and each core. If a sentences is being decoded on one CPU is switched to another, the CPU cache on the new CPU must be repopulated, slowing down decoding. We will therefore investigate binding threads to specific cores.

\subsection{Stack Configurations}

The most popular stack configuration for phrase-based models is coverage cardinality, that is, hypotheses that have translated the same number of source words are stored in the same stack. This is implemented in Pharaoh, Moses and Joshua.

%Since stack pruning occurs in each stack independently of each other, this affect the decoder's search errors.

However, there are alternatives to this configuration. ~\citep{Och:2001a} uses a single stack for all hypotheses, ~\citep{Brown:1993} uses coverage stacks (ie. one stack per unique coverage vector) while ~\citep{peitz2012jane} and ~\citep{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning. While useful, these prior works present only one particular stack configuration each.

~\citep{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.

We shall re-visit the question of stack configuration with a particular emphasis on how they can help improve the tradeoff between speed and translation quality. We will do so in the context of the cube-pruning algorithm, the algorithm that we will be using and which was not available to many of the earlier work.

\subsection{Phrase-Table Optimizations}

For any phrase-table table of a realistic size, memory and loading time constraints requires us to use a load-on-demand implementation. Moses has several which we can make use of, each with differing performance characteristics. Figure~\ref{fig:moses-phrase-tables-time} shows the decoding speed for the fastest two implementations.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding speed with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\cite{Bogoychev.probing.2016} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.

We propose two optimizations. Firstly, the translation rule caching mechanism in Moses saves the most recently used rules. However, this require locking and active management in clearing of old rules. The result is \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2877	& 2540 (-12\%) \\ \hline
\end{tabular}
\end{center}
\caption{Decoding speed (in words / sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}

We shall explore a simpler caching mechanism by creating a static cache of the most likely translation rules to be used at the start of decoding.

Secondly, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule. Compression was championed by ~\citep{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads. We shall therefore take the opposite approach to and improve decoding speed by disabling compression.

\subsection{Lexicalized Re-ordering Model Optimizations}

Similar to the phrase-table, the lexicalized re-ordering model is trained on parallel data. A resultant model file is then queried during decoding. The need for random lookup during querying inevitably results in slower decoding speed. Previous work such as~\citep{junczys_tsd_2012b} improve querying speed with more efficient data structures.

However, the model's query keys are the source and target phrase of each translation rule. Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file. However, the model remain the same under the log-linear framework, including having its own weights. %The search algorithm will produce identical results to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.

This optimization has precedent in ~\citep{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.

\section{Experimental Setup}
\label{sec:Experimental Setup}

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}

For verification with a different dataset, we also used a second system trained on the French-English Europarl corpus (2m parallel sentences). The two different systems have characterics that we are interested in analyzing; ar-en have short sentences with large models while fr-en have overly long sentences with smaller models. Where we need to compare model scores, we used held out test sets.

\begin{table}[h]
\begin{center}
\small
\begin{tabular}{|l|r|r|} \hline
		& ar-en		& fr-en \\ \hline
\multicolumn{3}{|l|}{For speed testing} \\ \hline
Set name	& \multicolumn{2}{|c|}{Subset of training data} \\
\# sentences  	& 800k 		& 200k \\
\# words 	& 5.8m 		& 5.9m \\ 
Avg words/sent	& 7.3		& 29.7 \\ \hline
\multicolumn{3}{|l|}{For model score testing} \\ \hline
Set name	& OpenSubtitles	& newstest2011 \\
\# sentences  	& 2000 		& 3003 \\
\# words 	& 14,620 	& 86,162 \\ 
Avg words/sent	& 7.3		& 28.7 \\ \hline
\end{tabular}
\end{center}
\caption{Test sets}
\label{tab:test-sets}
\end{table}

Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{the pop-limit was chosen from public discussion on the Moses mailing list on an acceptable balance between decoding speed and translation quality with Moses for commercial use %https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html 
}, rather than the basic phrase-based algorithm. The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.

As a baseline, we use a recent\footnote{The experiments were performed between January and May 2016 with the latest github code to hand. The main ar-en experiments were rerun with the source code as of 8th June, 2016 to ensure there were no material difference. The commit hash was \scriptsize{bc5f8d15c6ce4bc678ba992860bfd4be6719cee8} } version of the Moses decoder taken from the github repository.

For all experiments, we used a Dell PowerEdge R620 server with 16 cores, 32 hyper-threads, split over 2 physical processors (Intel Xeon E5-2650 @ 2.00GHz). The server has 380GB RAM. The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost 1.59\footnote{http://boost.org/} and the tcmalloc library.

\section{Results}
\label{sec:Results}

\subsection{Optimizing Memory}

Each sentence is decoded from start to finish using one worker thread. %The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.

As can be seen in Table~\ref{tab:optimizing-memory}, over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, dampening the scalability of the decoder. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
\begin{table}[h]
\begin{center}
\small
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

Figure~\ref{fig:mempool} compares the decoding speed for Moses and our decoder, using the same models, parameters and test set. Our decoder is over 3 times faster with one thread, and 4.7 times faster using all cores. Like Moses, performance actually worsens after approximately 15 threads, however, the problem is not as pronounced. This gives us a better foundation on which to build further innovations for fast, multi-core decoding.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{mempool.png}} 
\end{tabular}
\caption{Decoding time of Moses and our decoder, using the same models}
\label{fig:mempool}
\end{figure}

% A potential disadvantage of using memory pool is that the decoder will use more memory, however, Table~\ref{tab:memory-usage} shows that peak memory consumption are comparable.
% \begin{table}[h]
% \begin{center}
% \begin{tabular}{|l|c|c|} \hline
% 		& 1 thread	& 32 threads \\ \hline
% Moses  		& ???	 	& ??? \\
% Our work 	& ??? 		& ??? \\ \hline
% \end{tabular}
% \end{center}
% \caption{Peak Memory Usage in GB}
% \label{tab:memory-usage}
% \end{table}

\subsection{Stack Configuration}

We shall investigate the effects of the following three stack configurations on model score and decoding speed:
\begin{enumerate}
  \item \vspace{-2 mm} coverage cardinality,
  \item \vspace{-2 mm} coverage,
  \item \vspace{-2 mm} coverage and end position of most recently translated source word.
\end{enumerate}
Coverage cardinality is the same as that in Moses and Joshua. Coverage configuration uses one stack per unique coverage vector.

\emph{Coverage and end position of most recently translated source word} extends the coverage configuration by separating hypotheses where the last translate word are different. This is an optimization to reduce the number of checks on the distortion limit, which is dependent on the last word position.

The check is a binary function $d(C_{h}, e_{hypo}, range_r)$, where $C_{h}$ is the coverage vector of hypothesis $h$, $e_{h}$ is the end position of most recent source word that has been translated, and $range_r$ is the coverage of the rule to be applied. Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} shows the standard hypothesis expansion algorithm.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $h$ in $stack_{|C|}$} 
  \FORALL{$r$ in translation rules}
      \IF {$d(C_{h}, e_{hypo}, range_r)$}
	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
	\STATE add $h'$ to stack $stack_{C'}$
      \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Cardinality Stacks}
\label{algo:Hypothesis Expansion with Unsorted Stack}
\end{figure}

By grouping hypotheses according to coverage \emph{and} end position into groups we call \emph{'ministack'}, the distortion limit only needs to be checked for each group, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}. 
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
  \FORALL{$r$ in translation rules}
    \IF {$d(C_{h}, e_{hypo}, range_r)$}
      \FORALL{ $h$ in $ministack_{C, e}$} 
	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
	\STATE add $h'$ to $ministack_{C',e'}$
      \ENDFOR
    \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
\label{algo:Hypothesis Expansion with Sorted Stack}
\end{figure}

Also, stack pruning occurs on each hypothesis group independently, therefore, potentially affect search errors and model scores.
